### 2020-06-22 article.tex
WOrk on PCHPDDM section

On the other hand, the High-Performance unified framework for Domain Decomposition Methods (HPDDM)
is a much smaller project focusing on robust and scalable domain decomposition
preconditioners~\cite{jolivet2013scalable} and advanced iterative
methods~\cite{jolivet2016block} which can efficiently deal with linear
systems with multiple right-hand sides (also known and as block iterative methods)
and when there is a recurrence of varying coefficient matrices and right-hand
sides (recycling iterative methods). For a introduction to domain decomposition methods, see
~\cite{smith2004domain,dolean2014ddm,olofbook,tarekmathewbook} {\color{red} FIX CITATIONS}.

@@

{\color{red} TODO: move KSPMatSolve discussion out, and clean up the section}

@@

@@

Historically, the block Jacobi and overlapping Schwarz methods were implemented
as \pk{PCBJACOBI} and \pk{PCASM} respectively. ASM stands for additive Schwarz
method, as originally introduced by Schwarz~\cite{schwarz1870alternating}, but
the PETSc implementation is more versatile in the sense that one can easily
switch between this method and other variants such as the restricted additive
Schwarz method~\cite{cai1999restricted}. More details about overlapping domain decomposition
preconditioners will be given in~\cref{sec:interface-pc}, but first the other
related preconditioners will be quickly presented.

% In PETSc, there is also an implementation of two-level overlapping Schwarz
% methods using a generalized Dryja, Smith, Widlund
% preconditioner~\cite{dohrmann2008domain}. Its type is \pk{PCEXOTIC}.
When it comes to nonoverlapping preconditioners, \pk{PCNN} implements a basic balancing
Neumann--Neumann method~\cite{mandel1993balancing}. Much more advanced
techniques are available in \pk{PCBDDC}~\cite{zampini2016pcbddc}: adaptive selection
of primal constraints~\cite{pechstein2017unified}, change of basis for problems in
$H(\text{curl})$~\cite{dohrmann2016bddc}$\ldots$ For multigrid,
\pk{PCMG} exposes a machinery for geometric or algebraic preconditioners. In
particular, it is used for the implementation of the smoothed-aggregation
\pk{PCGAMG}~\cite{adams2004ultra}. There are also interfaces to other well-known
multigrid packages such as \pk{PCHYPRE} or \pk{PCML}~\cite{gee2006ml} from Trilinos. Of
course, there are many other available domain decomposition preconditioners,
either accessible through high-level libraries, e.g., FROSch~\cite{Heinlein:2016:PIT} in
Trilinos, or as stand-alone packages, e.g., BDDCML~\cite{vsistek2013parallel} or
FEMPAR-BDDC~\cite{badia2014highly}.

HPDDM implements the same one-level overlapping Schwarz methods as PETSc, albeit
emphasis is also put on optimized Schwarz methods~\cite{gander2006optimized}.
The main theoretical framework for defining robust two-level methods is called GenEO,
for generalized eigenvalue problem on the overlap~\cite{spillane2011robust,spillane2013abstract}.
GenEO was also implemented in other libraries~\cite{butler2020high}, but to the best of our knowledge,
the interface between HPDDM and PETSc proposes the first high-level access to
GenEO so that it can, for example, be used as a block-decomposed
solver~\cite{brown2012composable}.

  \subsection{Interfacing HPDDM overlapping Schwarz methods in PETSc\label{sec:interface-pc}}\
\pk{PCHPDDM}, the interface between HPDDM overlapping Schwarz methods and PETSc, is automatically
registered since PETSc version 3.12 when one configures PETSc with the extra
flag \pk{-{}-download-hpddm -{}-download-slepc}. It is then
possible to select the corresponding \pk{PC}
using the command line option \pk{-pc\_type hpddm}, or the routine \pk{PCSetType}(pc,
\pk{PCHPDDM}).% and its Fortran equivalent.
In this section, it is assumed that one wants to solve
    Ax = b,
for a given coefficient matrix $A$ and right-hand side $b$ of dimension $n$
assembled on $N$ processes, using a ksp($A$, $P_A$), i.e., the matrix used to
define the preconditioner is $A$ itself. Without anything more from the user,
\pk{PCHPDDM} is strictly equivalent to \pk{PCASM}. That is, if $A$ is distributed among
$N$ processes the preconditioner $M^{-1}$ is:
\begin{equation}\label{eq:one}
    M^{-1} = \sum_{i = 1}^N \tilde{R}_i^T (R_i A R_i^T)^{-1} R_i,
where $\{R_i\}_{i=1}^N$ are restriction operators that act on a global vector and
return a vector local to each process, possibly with some overlap.
$\{\tilde{R}_i\}_{i=1}^N$ are the same operators except that coefficients on the
overlap are set to 0. The action of the inverse of all $\{R_i A R_i^T\}_{i=1}^N$
can be parameterized through a local \pk{KSP}. If it is assumed that $A$ is the
discretization of a linear operator $\mathcal{L}$, e.g., $\mathcal{L}=-\Delta$, on a domain $\Omega$, it is
extremely common to exhibit parallelism by distributing $\Omega$ on $N$
processes, possibly with some overlap, and to construct the $\{R_i\}_{i=1}^N$ so
that they map global unknowns to unknowns local to each process. In addition to

@@

equation with homogeneous Dirichlet boundary conditions discretized using $\mathbb{P}_1$ finite elements, an explicit representation of these various operators is given~\cref{fig:poisson} for a

@@

Then, it is possible to enrich the original one-level preconditioner
from~\cref{eq:one} using a spectral coarse grid built in the following way:
    \item have SLEPc solve the local generalized eigenvalue problem concurrently:
        \begin{equation}\label{eq:gevp}
            \mathring{A}_i \Lambda_i = \lambda_i \tilde{R}_i R_i^T (R_i A R_i^T) \tilde{R}_i R_i^T \Lambda_i ;
    \item retrieve the $\nu_i$ smallest eigenpairs $\{\Lambda_{i_j}, \lambda_{i_j}\}_{j=1}^{\nu_i}$ and assemble a local deflation dense matrix $W_i = \begin{bmatrix} \tilde{R}_i R_i^T \Lambda_{i_1} & \cdots & \tilde{R}_i R_i^T \Lambda_{i_{\nu_j}} \end{bmatrix}$;
    \item define a global deflation matrix $P = \begin{bmatrix} R_1^T W_1 & \cdots & R_N^T W_N\end{bmatrix}$ and a new two-level preconditioner using the Galerkin product of $A$ and $P$:
        \begin{equation}\label{eq:two}
            M'^{-1}_{\text{additive}} = P \left(P^T A P\right)^{-1} P^T + \sum_{i = 1}^N \tilde{R}_i^T (R_i A R_i^T)^{-1} R_i.
Note that $\{\tilde{R}_i R_i^T\}_{i=1}^N$ define a partition of unity, i.e.,
    \sum_{i = 1}^N R_i^T \tilde{R}_i R_i^T R_i = I.
When used without PETSc and thus SLEPc, step \#1 is solved in HPDDM with
ARPACK~\cite{lehoucq1998arpack}. Step \#3 is common to most domain
decomposition or multigrid preconditioners. However, in HPDDM, the special
structure of the deflation matrix~$P$ which is sparse with dense blocks is
exploited. Interested readers are referred to the paper describing the technical
details of the implementation~\cite{jolivet2013scalable}. One important point to
notice is that the size of the coarse operator $P^TAP$ is much lower
than $n$, the size of $A$. Furthermore, coarsening is much more aggressive than
algebraic or geometric multigrids, see~\cref{sec:linear-elasticity} for
evidence. It is thus common to have the coarse operator distributed on
fewer processes than for $A$, see for example the concept of telescoping later
introduced in PETSc~\cite{may2016extreme}. In HPDDM, the coarse operator is
computed and redistributed simultaneously. This could be technical to implement
in PETSc as the result of the routine \pk{MatPtAP} is distributed on the same communicator as the one of the input matrix $A$.

% Of course, users do not have to implement the three previous steps themselves.
The auxiliary operators $\{\mathring{A}_i\}_{i=1}^N$ have simply to be provided
as well as the correspondence from local to global numbering,
see~\cref{fig:poisson}, using the routine \pk{PCHPDDMSetAuxiliaryMat}.  If there
is a \pk{DM} attached to the \pk{PC} of type \pk{DMPLEX}~\cite{knepley2009mesh}
or \pk{DMP4EST}~\cite{BursteddeWilcoxGhattas11}, then these auxiliary
overlapping operators and the correspondence are automatically assembled by
PETSc for users, see~\cref{fig:p4est}. Other \pk{DMType} may be supported by
implementing the proper \pk{DMCreateNeumannOverlap} callback, e.g., \pk{DMDA}
support could be easily added upon request. With the local
matrix pencils available, new options will be automatically registered such as
\pk{-pc\_hpddm\_levels\_1\_eps\_nev} or \pk{-pc\_hpddm\_levels\_1\_eps\_threshold}, which
may be used to set how many eigenpairs SLEPc should compute in step \#2.  The
default \pk{-eps\_nev} value is 20, i.e., the dimension of the coarse space is
$20\times N$ under the assumption that all concurrent SLEPc solves converge.
It is usually best to keep the same number of eigenvectors on each process for multiple reasons:
\begin{itemize}
    \item with standard eigensolvers such as ARPACK or \pk{EPSKRYLOVSCHUR}, it is not possible to compute all eigenpairs in a given interval of $\mathbb{R}^+$. Instead, an upper bound of the number of eigenpairs has to be provided, and then it is possible in \pk{PCHPDDM} to remove eigenvectors associated to eigenvalues above the given \pk{-eps\_threshold}; extra eigenvectors are thus computed, but immediately thrown away;
    \item if the number of eigenvectors is the same, coarse operators take advantage of block structures and are assembled as \pk{MatMPISBAIJ} or \pk{MatMPIBAIJ} matrices, with usually large block sizes, by default 20; this can have serious memory and performance implications, see~\cref{fig:perfAX};
    \item during solution phases, it is possible to use the functions \texttt{MPI\_Gather} and \texttt{MPI\_Scatter} to go from one to another level and vice versa, instead of the variable size alternatives \texttt{MPI\_Gatherv} and \texttt{MPI\_Scatterv}, which are more costly.
\end{itemize}
Another important option is \pk{-pc\_hpddm\_coarse\_p}, which sets the size of the
communicator on which the coarse operator should be distributed. The default
value is 1, i.e., the matrix $P^TAP$ is centralized on a single process.
The action of both inverses from~\cref{eq:two} can be parameterized through the
\pk{-pc\_hpddm\_coarse\_ksp\_type} and \pk{-pc\_hpddm\_levels\_1\_ksp\_type} options
respectively.
\begin{figure}[htbp]
  \centering
    \subfloat[Initial nonoverlapping decomposition.]{
        \includegraphics[width=0.45\textwidth]{Data/p4est_nonoverlapping.pdf}
    }
    \subfloat[Overlapping decomposition on which auxiliary operators $\{\mathring{A}_i\}_{i=1}^N$ are assembled.]{
        \includegraphics[width=0.45\textwidth]{Data/p4est_overlapping.pdf}
    }
    \caption{Automatic generation of overlapping subdomains when using a~\pk{DMP4EST}. For clarity, only four subdomains, one of which is disconnected, are colored.\label{fig:p4est}}
\end{figure}

    \subsubsection{Coarse corrections and eigenvalue problems}
Some other parameters may be adjusted to define how the one-level preconditioner
and the coarse grid operator interact with each other. In fact, the additive
correction from~\cref{eq:two} is the simplest to write down, but most often not
very efficient numerically~\cite{tang2009comparison}, and thus not used as
default in \pk{PCHPDDM}. This observation is not specific to domain
decomposition methods and also applies to multigrid~\cite{bastian1998additive}.
Users can choose between $M'^{-1}_{\text{additive}}$ and:
\begin{align*}
    M'^{-1}_{\text{deflated}} &= Q + M^{-1}\left(I - AQ\right)\quad
    \text{(default)}\\
    M'^{-1}_{\text{balanced}} &= Q + \left(I - AQ\right)M^{-1}\left(I - AQ\right),
\end{align*}
with $Q = P \left(P^T A P\right)^{-1} P^T$. These preconditioners may be selected using the
options \pk{-pc\_hpddm\_coarse\_correction} \pk{(deflated$|$additive$|$balanced)} or the
routine \pk{PCHPDDMSetCoarseCorrectionType}(pc, type) and its Fortran equivalent.

The local generalized eigenvalue problems~\cref{eq:gevp} solved by SLEPc can also be adjusted.
For example, in addition to the local unassembled Neumann matrices~$\{\mathring{A}_i\}_{i=1}^N$,
if one has access to local matrices with optimized
boundary conditions $\{B_i\}_{i=1}^N$, e.g., Robin transmission conditions, it
is possible to pass these using the routine \pk{PCHPDDMSetRHSMat}(pc, $B_i$).
The following generalized eigenvalue problems, called GenEO-2 in the
literature~\cite{haferssas2017soras}, are then solved in each subdomain:
    \mathring{A}_i \Lambda_i = \lambda_i B_i \Lambda_i.
    \subsubsection{Multilevel extension}\label{sec:ml}
In the previous sections, it was shown how to supply the required information to
\pk{PCHPDDM} in order to build robust two-level overlapping domain decomposition preconditioners.
A multilevel extension of the GenEO framework is also
possible~\cite{aldaas2019multi}. It can be used without any additional
information provided by users. Since the dimension of the
coarse operator of the second level grows with the number of subdomains of the
fine level, switching to a multilevel scheme may alleviate the increasing cost
of solving coarse systems. \pk{PCHPDDM} follows the same numbering of
\pk{PCBDDC}%\footnote{which itself follows the opposite numbering of \pk{PCMG}}
: the finest level is always numbered 1, the level index increases, up until
the coarsest level whose options are always prefixed by \pk{coarse\_}. In order to
register an additional level $l' = l+1$, the following conditions must be met at
level $l$:
    \item there must be more than one subdomain;
    \item at least one local eigenvector must be computed.
Here are some examples with $N = 16$ subdomains on the finest level:
    \item \pk{-pc\_hpddm\_levels\_1\_eps\_threshold 0.4 -pc\_hpddm\_coarse\_p 8} would
        define a two-level method with the coarse operator being distributed
        among 8 processes, assuming that there is globally enough $\lambda_i$ in~\cref{eq:gevp} lower than 0.4;
    \item \pk{-pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_p 8} \linebreak
        \pk{-pc\_hpddm\_levels\_2\_eps\_nev 10 -pc\_hpddm\_coarse\_p 2} would define
        a three-level method with the second level (resp.~ coarse operator)
        being distributed among 8 (resp.~2) processes;
    \item \pk{-pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_p 8} is not
        valid because there are only two levels, so the options \pk{levels\_2\_} are
        not registered and coarse\_ has to be used instead;
    \item \pk{-pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_eps\_nev 10} \linebreak \pk{-pc\_hpddm\_coarse\_p 2}
        is not valid because there is no option \pk{levels\_2\_p}, so the default
        value of 1 is used for the second level, and it is thus not possible to
        define a third level with a single second-level subdomain.
    \subsubsection{MatIS}

@@

finite element context. Using a local to global mapping and \pk{MatConvert}, it
is possible to reconstruct the local assembled matrices, equivalent to the

@@

fact no overlap. It is common in the field of BDD to see the coarse problem as
an unassembled operator as well. In fact, this is mandatory to define efficient
multilevel variants, as implemented in \pk{PCBDDC}. In~\cref{sec:bddc}, it will
be shown how a coarse problem obtained in \pk{PCBDDC} may be solved using
\pk{PCHPDDM}.

@@

Considering only small displacements, this linear problem with highly heterogeneous elastic moduli is solved in 3D.

@@

$(\mathbb{P}_1, \mathbb{P}_1, \mathbb{P}_1)$ finite elements.

@@

The number of unknowns is $\pgfmathprintnumber{593} \times 10^6$, with approximately $\pgfmathprintnumber{45}$ nonzero coefficients per row.
There are heterogeneities due to jumps in $E$ and $\nu$. The following discontinuous piecewise constant values are considered: $(E_1, \nu_1) = (2\times 10^{11}, 0.25)$, $(E_2, \nu_2) = (10^{7}, 0.45)$, see~\cref{fig:elas}.

@@

    \item \pk{PCHPDDM} with an inexact coarse solver using \pk{PCHYPRE};\label{enum:inexacthypre}
    \item \pk{PCHPDDM} with inexact coarse solvers using GenEO multilevel extension, cf.~\cref{sec:ml}.\label{enum:inexact}

@@

The options \pk{-ksp\_gmres\_modifiedgramschmidt

@@

-ksp\_type fgmres} are used in conjunction with all \pk{PC}. The options specific to each solver described above are given next. \\[-4pt]

@@

By default, the coarse problem in \pk{PCHPDDM} is solved using an exact $LU$ or Cholesky factorization, in this case using 24 processes and MKL CPARDISO. Switching to an inexact solver using \pk{PCHYPRE} (resp.~GenEO multilevel extension), cf.~\cref{enum:inexacthypre} (resp.~\cref{enum:inexact}), is straightforward. \\[-4pt]

@@

$M$ is the number of subdomains used to define the second-level domain decomposition, which is an aggregation of first-level subdomain.

@@

it is 1.015. The code used to produce these results was borrowed from

>>>>

On the other hand, the High-Performance unified framework for Domain Decomposition Methods (HPDDM)
is a much smaller project focusing on robust and scalable domain decomposition
preconditioners~\cite{jolivet2013scalable} and advanced iterative
methods~\cite{jolivet2016block} which can efficiently deal with linear
systems with multiple right-hand sides (also known and as block iterative methods)
and when there is a recurrence of varying coefficient matrices and right-hand
sides (recycling iterative methods). For an introduction to domain decomposition methods, see
~\cite{smith2004domain,dolean2014ddm,olofbook,tarekmathewbook} {\color{red} FIX CITATIONS}.

@@

{\color{red} TODO: move KSPMatSolve discussion out, and clean up the section, add results for irk?}

@@

For an introduction to domain decomposition methods, see
~\cite{smith2004domain,dolean2014ddm,olofbook,tarekmathewbook} {\color{red} FIX CITATIONS}.

@@

Basic one-level methods such as the block Jacobi and the (overlapping) additive Schwarz method are implemented
in PETSc as \pk{PCBJACOBI} and \pk{PCASM} respectively, with some additional customization~\cite{cai1999restricted}
available for \pk{PCASM}. However, none of these methods provide automatic support for a coarse level solver,
which is mandatory to obtain algorithmic scalability with a large number of processes.
%ASM stands for additive Schwarz
%method, as originally introduced by Schwarz~\cite{schwarz1870alternating}, but
%the PETSc implementation is more versatile in the sense that one can easily
%switch between this method and other variants such as the restricted additive
%Schwarz method~\cite{cai1999restricted}. More details about overlapping domain decomposition
%preconditioners will be given in~\cref{sec:interface-pc}, but first the other
%related preconditioners will be quickly presented.
% In PETSc, there is also an implementation of two-level overlapping Schwarz
% methods using a generalized Dryja, Smith, Widlund
% preconditioner~\cite{dohrmann2008domain}. Its type is \pk{PCEXOTIC}.
Furthermore, PETSc provides domain decomposition methods of the non-overlapping type;
\pk{PCNN} implements a basic balancing
Neumann--Neumann method~\cite{mandel1993balancing}, while more advanced
preconditioning techniques are available with \pk{PCBDDC}~\cite{zampini2016pcbddc}, including guaranteed robust convergence
by means of adaptive selection
of primal constraints~\cite{pechstein2017unified} and support for $H(\text{curl})$ and $H(\text{div})$
conforming finite elements~\cite{zampinihcurl1,zampinihcurl2,zampinihdiv,zampinidarcy}, as well as for IsoGeometric Analysys \cite{zampiniiga1,zampiniiga2,zampiniiga3}.
For multigrid,
\pk{PCMG} provides a unified framework for geometric or algebraic preconditioners, which is used for the implementation of the smoothed-aggregation
\pk{PCGAMG}~\cite{adams2004ultra}. There are also interfaces to other well-known
multigrid packages such as \pk{PCHYPRE}~\cite{falgout2002hypre} or \pk{PCML}~\cite{gee2006ml} from Trilinos. Of
course, there are many other available domain decomposition preconditioners,
either accessible through high-level libraries, e.g., FROSch~\cite{Heinlein:2016:PIT} in
Trilinos, or as stand-alone packages, e.g., BDDCML~\cite{vsistek2013parallel} or
FEMPAR-BDDC~\cite{badia2014highly}.

HPDDM implements the same one-level overlapping Schwarz methods as PETSc as well as
optimized Schwarz methods~\cite{gander2006optimized} {\color{red} Are you planning to expose them too?}.
It also provides support for defining robust two-level methods using the GENeralized Eigenvalue problem on the Overlap (GenEO) framework~\cite{spillane2011robust,spillane2013abstract}; the results presented in this work presents the first high-level access to
GenEO from PETSc and its composable solver infrastructrure~\cite{brown2012composable}.
We note that GenEO has been recently implemented in other libraries as well~\cite{butler2020high}

  \subsection{Interfacing HPDDM overlapping Schwarz methods in PETSc\label{sec:interface-pc}}\
The interface between HPDDM overlapping Schwarz methods and PETSc, \pk{PCHPDDM}, is automatically
registered since PETSc version 3.12 when one configures PETSc with the extra
flag \pk{-{}-download-hpddm -{}-download-slepc}. It is then
possible to select the corresponding \pk{PC}
using the command line option \pk{-pc\_type hpddm}, or the routine \pk{PCSetType}(pc,
\pk{PCHPDDM}).
 % and its Fortran equivalent.
In this section, we are interested in solving
\begin{equation}
    Ax = b,
\end{equation}
for a given coefficient matrix $A$ and right-hand side $b$ of dimension $n$
assembled on $N$ processes, using a ksp($A$, $P_A$), i.e., the matrix used to
define the preconditioner is $A$ itself {\color{red} WHY IS THAT? You do not support a different preconditioning matrix?}. Without anything more from the user,
\pk{PCHPDDM} is strictly equivalent to \pk{PCASM}. That is, if $A$ is distributed among
$N$ processes, the action of the preconditioner $M^{-1}$ is:
\begin{equation}\label{eq:one}
    M^{-1} = \sum_{i = 1}^N \tilde{R}_i^T (R_i A R_i^T)^{-1} R_i,
\end{equation}
where $\{R_i\}_{i=1}^N$ are restriction operators that act on a global vector and
return a local vector on each process, possibly with some overlap.
$\{\tilde{R}_i\}_{i=1}^N$ are the same operators except that coefficients on the
overlap are set to 0~\cite{cai1999restricted}. The action of each $\{R_i A R_i^T\}^{-1}_{i}$
can be parameterized through a local \pk{KSP}. When $A$ is the
discretization of a linear partial differential operator operator $\mathcal{L}$ on a domain $\Omega$, it is
extremely common to exhibit parallelism by distributing $\Omega$ on $N$
processes, possibly with some overlap, and to construct the $\{R_i\}_{i=1}^N$ so
that they map global unknowns to unknowns local to each process.

In addition to

@@

equation with homogeneous Dirichlet boundary conditions discretized using $\mathbb{P}_1$ finite elements, an explicit representation of these  operators is given~\cref{fig:poisson} for a

@@

With extra Neumann information at hand, it is then possible to enrich the original one-level preconditioner
from~\cref{eq:one} using a spectral coarse grid built in the following way:
\begin{enumerate}
    \item have SLEPc solve the local generalized eigenvalue problem concurrently:
        \begin{equation}\label{eq:gevp}
            \mathring{A}_i \Lambda_i = \lambda_i \tilde{R}_i R_i^T (R_i A R_i^T) \tilde{R}_i R_i^T \Lambda_i ;
        \end{equation}
    \item retrieve the $\nu_i$ smallest eigenpairs $\{\Lambda_{i_j}, \lambda_{i_j}\}_{j=1}^{\nu_i}$ and assemble a local deflation dense matrix $W_i = \begin{bmatrix} \tilde{R}_i R_i^T \Lambda_{i_1} & \cdots & \tilde{R}_i R_i^T \Lambda_{i_{\nu_j}} \end{bmatrix}$;
    \item define a global deflation matrix $P = \begin{bmatrix} R_1^T W_1 & \cdots & R_N^T W_N\end{bmatrix}$ and a new two-level preconditioner using the Galerkin product of $A$ and $P$:
        \begin{equation}\label{eq:two}
            M'^{-1}_{\text{additive}} = P \left(P^T A P\right)^{-1} P^T + \sum_{i = 1}^N \tilde{R}_i^T (R_i A R_i^T)^{-1} R_i.
        \end{equation}
\end{enumerate}
Note that $\{\tilde{R}_i R_i^T\}_{i=1}^N$ defines a partition of unity, i.e.,
\begin{equation*}
    \sum_{i = 1}^N R_i^T \tilde{R}_i R_i^T R_i = I.
\end{equation*}
{\color{red} Should we mention the connections with the condition number here? and thus discuss the guaranteed convergence property of the method?}

%When used without PETSc and thus SLEPc, step \#1 is solved in HPDDM with
%ARPACK~\cite{lehoucq1998arpack}.
While Step \#3 is common to most domain
decomposition or multigrid preconditioners, HPDDM takes advantage of the special
block sparse structure of the deflation matrix~$P$.
Moreover,
the size of the coarse operator $P^TAP$ is much lower
than $n$, the size of $A$, and coarsening is much more aggressive than
when using algebraic or geometric multigrids methods.
%, see~\cref{sec:linear-elasticity} for
%evidence.
It is thus natural to remap the coarse operator on
a subset of the processes used to decompose the original matrix $A$; in HPDDM, the coarse operator is
computed and redistributed simultaneously.
Interested readers are referred to the paper describing the technical
details of the implementation~\cite{jolivet2013scalable}.
We finally note here that similar techniques have been proven beneficial for other types
of solvers in PETSc~\cite{may2016extreme}.

% Of course, users do not have to implement the three previous steps themselves.
In order to use GenEO from PETSc, the auxiliary operators $\{\mathring{A}_i\}_{i=1}^N$
and the correspondending local to global map of degrees of freedom (see~\cref{fig:poisson})
can be provided using the routine \pk{PCHPDDMSetAuxiliaryMat}.
In special cases where the matrix has been constructed using the PETSc discretization
instrastucture, the relevant information is automatically computed
by PETSc. In \cref{fig:p4est} we provide an example of non-overlapping (left) and overlapping (right)
decomposition using a non-conforming grid with \pk{DMPLEX}~\cite{knepley2009mesh}. The
mesh has been constructed internally by \pk{DMPLEX} using a random pattern of non-conforming 2:1 refinements
and a space-filling curve distribution by p4est~\cite{BursteddeWilcoxGhattas11}.
The required elements needed to properly assemble the Neumann operators in a conforming way are pictured with a lighter color.
Other \pk{DMType}s, including user-defined, may be supported by
implementing the proper \pk{DMCreateNeumannOverlap} callback; support for \pk{DMDA}
could be easily added upon request.
\begin{figure}[htbp]
  \centering
    \subfloat[Initial nonoverlapping decomposition.]{
        \includegraphics[width=0.45\textwidth]{Data/p4est_nonoverlapping.pdf}
    }
    \subfloat[Overlapping decomposition on which auxiliary operators $\{\mathring{A}_i\}_{i=1}^N$ are assembled.]{
        \includegraphics[width=0.45\textwidth]{Data/p4est_overlapping.pdf}
    }
    \caption{Automatic generation of overlapping subdomains when using a~\pk{DMP4EST}. For clarity, only four subdomains, one of which is disconnected, are colored.\label{fig:p4est}}
\end{figure}

With the local
matrix pencils available, new options will be automatically registered such as
\pk{-pc\_hpddm\_levels\_1\_eps\_nev} or \pk{-pc\_hpddm\_levels\_1\_eps\_threshold}, which
may be used to customize the eigenpairs SLEPc will compute in step \#2.  The
default \pk{-eps\_nev} value is 20, i.e., the dimension of the coarse space is
$20\times N$ under the assumption that all concurrent SLEPc solves converge.

For performance reasons, it is usually better to compute the same number of eigenvectors
on each process, since in this case the coarse operator will be assembled using the \pk{MatMPISBAIJ} (or \pk{MatMPIBAIJ}) matrix format,
with usually large block sizes; this can have serious memory and performance implications, see~\cref{fig:perfAX}. During solution phases,
we can use \texttt{MPI\_Gather} and \texttt{MPI\_Scatter} operations to move between levels,
instead of the variable size alternatives \texttt{MPI\_Gatherv} and \texttt{MPI\_Scatterv} that could exhibit worse performances.
In addition, standard eigensolvers such as ARPACK~\cite{lehoucq1998arpack} or \pk{EPSKRYLOVSCHUR} from SLEPc,
cannot compute all eigenpairs in a given interval of $\mathbb{R}^+$. Instead, an upper bound of the number of eigenpairs has to be provided.

The option \pk{-pc\_hpddm\_coarse\_p} can be used to provide the size of the
subcommunicator that will be used to remap the coarse operator. With the default
value of 1, the coarse matrix $P^TAP$ is centralized on a single process.
The action of the inverses in~\cref{eq:two} can be parameterized through the
\pk{-pc\_hpddm\_coarse\_ksp\_type} and \pk{-pc\_hpddm\_levels\_1\_ksp\_type} options
respectively.

    \subsubsection{Coarse corrections customization}
Some other parameters may be adjusted to define how the one-level preconditioner
and the coarse grid operator interact. Besided simple the additive
correction shown in~\cref{eq:two}, users can choose more numerically efficient
corrections as:~\cite{tang2009comparison,bastian1998additive}
\begin{align*}
    M'^{-1}_{\text{deflated}} &= Q + M^{-1}\left(I - AQ\right)\quad
    \text{(default)}\\
    M'^{-1}_{\text{balanced}} &= Q + \left(I - AQ\right)M^{-1}\left(I - AQ\right),
\end{align*}
with $Q = P \left(P^T A P\right)^{-1} P^T$. These preconditioners may be selected using the
options \pk{-pc\_hpddm\_coarse\_correction} \pk{(deflated$|$additive$|$balanced)} or the
routine \pk{PCHPDDMSetCoarseCorrectionType}.

    \subsubsection{Extension of the GenEO framework}
The local generalized eigenvalue problems~\cref{eq:gevp} can also be adjusted.
In addition to the local unassembled Neumann matrices~$\{\mathring{A}_i\}_{i=1}^N$,
users can prescribe additional local operators defined on the overlapped subdomain and
with optimized, Robin
boundary conditions $\{B_i\}_{i=1}^N$ via
the routine \pk{PCHPDDMSetRHSMat}(pc, $B_i$).
In such a case, the following generalized eigenvalue problems, called GenEO-2 in the
literature~\cite{haferssas2017soras}, are then solved in each subdomain:
\begin{equation*}
    \mathring{A}_i \Lambda_i = \lambda_i B_i \Lambda_i.
\end{equation*}
{\color{red} One sentence or two on why this is needed? what are the cases where this is used in practice?}

    \subsubsection{Multilevel extension}\label{sec:ml}
In the previous sections, it was shown how to supply the required information to
\pk{PCHPDDM} in order to build robust two-level overlapping domain decomposition preconditioners.
Since the dimension of the
coarse operator linearly depends on the number of subdomains of the
fine level decomposition, switching to a multilevel scheme may alleviate the increasing cost
of solving the coarse linear system. A multilevel extension of the GenEO framework has been recently
proposed~\cite{aldaas2019multi}, with the advantage that the multilevel hierarchy can be
automatically constructed without any additional information from the user.

\pk{PCHPDDM} follows the same numbering of \pk{PCBDDC}:
the finest level is always numbered with 1, and the level index increases as traversing the hierarchy, up until
the coarsest level, whose solver options are prefixed by \pk{coarse\_}. In order to
register an additional level $l' = l+1$, the following conditions must be met at
level $l$:
\begin{itemize}
    \item there must be more than one subdomain;
    \item at least one local eigenvector must be computed per coarse subdomain.
\end{itemize}
For example, using $N = 16$ subdomains on the finest level, the options

\pk{-pc\_hpddm\_levels\_1\_eps\_threshold 0.4 -pc\_hpddm\_coarse\_p 8}

\noindent will define a two-level method with the coarse operator being distributed
among 8 processes, assuming that there is globally enough $\lambda_i$ in~\cref{eq:gevp} smaller than the prescribed threshold 0.4.
Similarly, the options

\pk{-pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_p 8}

\pk{-pc\_hpddm\_levels\_2\_eps\_nev 10 -pc\_hpddm\_coarse\_p 2}

\noindent will define
a three-level method with the second level distributed among 8 processes, while the coarsest level
will be built from 10 eigenvalues from the second level and remapped onto 2 processes.
Additional examples for the solver customization are provided in~\cref{sec:pchpddm-examples}
%    \item \pk{-pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_p 8} is not
%        valid because there are only two levels, so the options \pk{levels\_2\_} are
%        not registered and coarse\_ has to be used instead;
%    \item \pk{-pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_eps\_nev 10} \linebreak \pk{-pc\_hpddm\_coarse\_p 2}
%        is not valid because there is no option \pk{levels\_2\_p}, so the default
%        value of 1 is used for the second level, and it is thus not possible to
%        define a third level with a single second-level subdomain.
%\end{itemize}

    \subsubsection{PCHPDDM for non-overlapping domain decomposition}

@@

finite element context. Since the assembled form of these operators can be reconstructed
it is possible to construct the Dirichlet operators, e.g.

@@

fact no overlap. {\color{red} add one sentence with reference to Nicole's work on GenEO for BDD?}
%It is common in the field of BDD to see the coarse problem as
%an unassembled operator as well. In fact, this is mandatory to define efficient
%multilevel variants, as implemented in \pk{PCBDDC}. In~\cref{sec:bddc}, it will
%be shown how a coarse problem obtained in \pk{PCBDDC} may be solved using
%\pk{PCHPDDM}.
  \subsection{Applications and numerical results}\label{sec:pchpddm-examples}

@@

We first report on solving the three-dimensional system of linear elasticy with highly heterogeneous elastic modulus.

@@

$(\mathbb{P}_1, \mathbb{P}_1, \mathbb{P}_1)$ finite elements resulting in $\pgfmathprintnumber{593} \times 10^6$ number of uknowns
with approximately $\pgfmathprintnumber{45}$ nonzero coefficients per row.

@@

There are heterogeneities due to jumps in $E$ and $\nu$. The following discontinuous piecewise constant values are considered: $(E_1, \nu_1) = (2\times 10^{11}, 0.25)$ in blue regions, while $(E_2, \nu_2) = (10^{7}, 0.45)$ in red regions, see~\cref{fig:elas}. {\color{red} Pierre, check the blue/red here and in the figure caption}
The code used to produce these results was borrowed from
GenEO~\cite{aldaas2019multi} and it is available

@@

    \item \pk{PCHPDDM} with an inexact coarse solver using \pk{PCHYPRE};\label{enum:inexacthypre} {\color{red} I think we should not include this case}
    \item \pk{PCHPDDM} with inexact coarse solvers using GenEO multilevel extension.\label{enum:inexact} %, cf.~\cref{sec:ml}.\label{enum:inexact}

@@

As a linear solver, we use flexible GMRES, customized with the options \pk{-ksp\_gmres\_modifiedgramschmidt

@@

-ksp\_type fgmres}. The options specific to each solver described above are given below. \\[-4pt]

@@

By default, the coarse problem in \pk{PCHPDDM} is solved using an exact $LU$ or Cholesky factorization, in this case using 24 processes and \pk{MKL\_CPARDISO}. Switching to an inexact solver using \pk{PCHYPRE} (resp.~GenEO multilevel extension), cf.~\cref{enum:inexacthypre} (resp.~\cref{enum:inexact}), is straightforward. \\[-4pt]

@@

For the multivel extension of GenEO, $M$ is the number of subdomains used to define the second-level domain decomposition, which is an aggregation of first-level subdomain.

@@

it is 1.015.
